---
title: "Responsibility for influencing others"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
  pdf_document:
    toc: yes
    toc_depth: 3
---
# Setup
```{r setup, include = FALSE, message = FALSE, warning = FALSE}
require(knitr)
library(ggstatsplot)
library(DT)
library(scales)
library(patchwork)
library(jpeg)
library(ggtext)
library(tidyverse)
library(brms)
library(cowplot)
library(Hmisc)
library(emmeans)
library(tidybayes)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

theme_set(theme_classic() + theme(text = element_text(size = 14)))

options(dplyr.summarise.inform = FALSE)
options(xtable.floating = FALSE)
options(xtable.timestamp = "")

set.seed(1)

load_exp_data <- function(data_dir, docs_dir = NULL) {
  trials_path <- file.path(data_dir, "trials.csv")
  stopifnot(file.exists(trials_path))

  df <- read_csv(trials_path, show_col_types = FALSE)

  if (!is.null(docs_dir)) {
    info_path <- file.path(docs_dir, "trial_info.csv")
    if (file.exists(info_path)) {
      trial_info <- read_csv(info_path, show_col_types = FALSE)

      join_keys <- intersect(c("trial", "name"), intersect(names(df), names(trial_info)))
      if (length(join_keys) > 0) {
        df <- df %>% left_join(trial_info, by = join_keys)
      }
    }
  }

  df
}


out_dir <- "../figures"

paths_exp1 <- list(
  data_dir    = "../data/experiment1",
  figures_dir = "../figures/experiment1",
  docs_dir    = "../docs/experiment1"
)

paths_exp2 <- list(
  data_dir    = "../data/experiment2",
  figures_dir = "../figures/experiment2",
  docs_dir    = "../docs/experiment2"
)

paths_exp3 <- list(
  data_dir    = "../data/experiment3",
  figures_dir = "../figures/experiment3",
  docs_dir    = "../docs/experiment3"
)

response_data_exp1 <- load_exp_data(paths_exp1$data_dir, paths_exp1$docs_dir)
response_data_exp2 <- load_exp_data(paths_exp2$data_dir, paths_exp2$docs_dir)
response_data_exp3 <- load_exp_data(paths_exp3$data_dir, paths_exp3$docs_dir)

```

# Responsibility judgments by role
```{r}
ROLE_BREAKS <- c(
  "root (with knowledge)",
  "root (without knowledge)",
  "dependent",
  "independent"
)

ROLE_COLORS <- c(
  "root (with knowledge)"    = "#9C7BD9",
  "root (without knowledge)" = "#C6B5EE",
  "dependent"                = "#F7C948",
  "independent"              = "#4FC3F7"
)


make_role_barplot_raw <- function(df,
                                  title = "Mean Responsibility Judgments",
                                  subtitle = NULL,
                                  subtitle_hjust = 0.5,
                                  y_max = 10.5) {

  df <- df %>% filter(!is.na(role_lbl), !is.na(response))

  ggplot(df, aes(x = role_lbl, y = response, fill = role_lbl)) +
    stat_summary(fun = mean, geom = "bar", width = 0.75) +
    stat_summary(fun.data = mean_cl_boot, geom = "linerange",
                 color = "black", linewidth = 2) +
    scale_fill_manual(
      values = ROLE_COLORS,
      breaks = ROLE_BREAKS,
      drop = TRUE
    ) +
    scale_x_discrete(drop = TRUE) +
    guides(
      fill = guide_legend(
        ncol = 4,
        byrow = TRUE,
        keywidth  = unit(1.2, "cm"),
        keyheight = unit(1.2, "cm")
      )
    ) +
    scale_y_continuous(
      limits = c(0, y_max),
      breaks = seq(0, y_max, by = 1),
      expand = c(0, 0)
    ) +
    labs(
      title = title,
      subtitle = subtitle,
      y = "Responsibility"
    ) +
    theme_minimal(base_size = 20) +
    theme(
      legend.position = "bottom",
      legend.spacing.x = unit(0.8, "cm"),
      legend.text = element_text(
        size = 28,
        margin = margin(l = 0.4, r = 2, unit = "cm")
      ),
      legend.title = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank(),
      axis.line = element_line(color = "black", linewidth = 0.2),
      axis.ticks = element_line(color = "black"),
      axis.ticks.length = unit(0.2, "cm"),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(color = "grey80", linewidth = 0.5),
      panel.grid.minor.y = element_blank(),
      axis.text.y = element_text(size = 25),
      axis.title.y = element_text(size = 25),
      plot.title = element_text(size = 32, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 28,color = "grey35",hjust = subtitle_hjust)
    )
}


make_plot_exp1 <- function(response_data) {
  df <- response_data %>%
    filter(!is.na(role), !is.na(response)) %>%
    mutate(
      role_lbl = case_when(
        role == "root" & knows == TRUE  ~ "root (with knowledge)",
        role == "root" & knows == FALSE ~ "root (without knowledge)",
        role == "dependent"             ~ "dependent",
        role == "independent"           ~ "independent",
        TRUE ~ NA_character_
      ),
      role_lbl = factor(role_lbl, levels = ROLE_BREAKS)
    )

  make_role_barplot_raw(
    df,
    title = "A. Experiment 1",
    subtitle = "Visible voting",
    subtitle_hjust = 0.51
  )
}

make_plot_exp2 <- function(response_data) {
  df <- response_data %>%
    filter(!is.na(role), !is.na(response)) %>%
    mutate(
      role_lbl = case_when(
        role == "root"        ~ "root (without knowledge)",
        role == "dependent"   ~ "dependent",
        role == "independent" ~ "independent",
        TRUE ~ NA_character_
      ),
      role_lbl = factor(role_lbl, levels = ROLE_BREAKS)
    )

  make_role_barplot_raw(
    df,
    title = "B. Experiment 2",
    subtitle = "Mechanical system",
    subtitle_hjust = 0.68
  )
}

make_plot_exp3 <- function(response_data) {
  df <- response_data %>%
    filter(!is.na(role), !is.na(response)) %>%
    mutate(
      role_lbl = case_when(
        role == "mentor"      ~ "root (with knowledge)",
        role == "mentee"      ~ "dependent",
        role == "independent" ~ "independent",
        TRUE ~ NA_character_
      ),
      role_lbl = factor(role_lbl, levels = ROLE_BREAKS)
    )

  make_role_barplot_raw(
    df,
    title = "C. Experiment 3",
    subtitle = "Advised voting",
    subtitle_hjust = 0.545
  )
}


p1 <- make_plot_exp1(response_data_exp1) +
  theme(
    legend.position = "bottom",
    legend.justification = c(-0.3, 0)
  )
p2 <- make_plot_exp2(response_data_exp2) + theme(legend.position = "none")
p3 <- make_plot_exp3(response_data_exp3) + theme(legend.position = "none")

patched <- (p1 | p2 | p3) +
  plot_layout(widths = c(4, 3, 3))

patched
```
```{r}
ggsave(
  filename = file.path(out_dir, "results.pdf"),
  plot = patched,
  width = 25,
  height = 8,
  dpi = 300
)
```
# Responsibility for root causes by follower proportion
```{r}
PROP_LEVELS <- c("0%", "50%", "100%")

PROP_COLORS <- c(
  "0%"   = "#FFF7CC",
  "50%"  = "#FFD84D",
  "100%" = "#F4A261"
)

as_num_safely <- function(x) {
  suppressWarnings(as.numeric(gsub("%", "", as.character(x))))
}

df_e1 <- response_data_exp1 %>%
  filter(role == "root", !is.na(response), !is.na(follow_proportion)) %>%
  mutate(
    fp = as_num_safely(follow_proportion),
    prop_lbl = case_when(
      fp == 0   ~ "0%",
      fp == 0.5 ~ "50%",
      fp == 1   ~ "100%",
      TRUE ~ NA_character_
    ),
    exp = "A. Experiment 1"
  ) %>%
  filter(!is.na(prop_lbl)) %>%
  transmute(exp, prop_lbl, response)

df_e2 <- response_data_exp2 %>%
  filter(role == "root", !is.na(response), !is.na(follow_proportion)) %>%
  mutate(
    fp = as_num_safely(follow_proportion),
    prop_lbl = case_when(
      fp == 0   ~ "0%",
      fp == 0.5 ~ "50%",
      fp == 1   ~ "100%",
      TRUE ~ NA_character_
    ),
    exp = "B. Experiment 2"
  ) %>%
  filter(!is.na(prop_lbl)) %>%
  transmute(exp, prop_lbl, response)

df_e3 <- response_data_exp3 %>%
  filter(role == "mentor", !is.na(response), !is.na(follow_proportion)) %>%
  mutate(
    fp = as_num_safely(follow_proportion),
    prop_lbl = case_when(
      fp == 0   ~ "0%",
      fp == 50  ~ "50%",
      fp == 100 ~ "100%",
      TRUE ~ NA_character_
    ),
    exp = "C. Experiment 3"
  ) %>%
  filter(!is.na(prop_lbl)) %>%
  transmute(exp, prop_lbl, response)


make_prop_barplot_raw <- function(df,
                                   title,
                                   subtitle = NULL,
                                   subtitle_hjust = 0.5,
                                   show_legend = FALSE) {

  df <- df %>%
    mutate(prop_lbl = factor(prop_lbl, levels = PROP_LEVELS)) %>%
    filter(!is.na(prop_lbl), !is.na(response))

  prop_counts <- df %>% count(prop_lbl, name = "n_trials")

  p <- ggplot(df, aes(x = prop_lbl, y = response, fill = prop_lbl)) +
    stat_summary(fun = mean, geom = "bar", width = 0.75) +
    stat_summary(
      fun.data = mean_cl_boot,
      geom = "linerange",
      color = "black",
      linewidth = 2
    ) +
    geom_text(
      data = prop_counts,
      aes(x = prop_lbl, y = 10.5, label = paste0("n=", n_trials)),
      inherit.aes = FALSE,
      vjust = -0.4,
      size = 5
    ) +
    scale_fill_manual(
      values = PROP_COLORS,
      breaks = PROP_LEVELS,
      limits = PROP_LEVELS,
      drop = TRUE
    ) +
    scale_x_discrete(drop = TRUE) +
    guides(
      fill = guide_legend(
        ncol = 3,
        byrow = TRUE,
        keywidth  = unit(1.2, "cm"),
        keyheight = unit(1.2, "cm")
      )
    ) +
    scale_y_continuous(
      limits = c(0, 10.5),
      breaks = seq(0, 10.5, by = 1),
      minor_breaks = seq(0, 10.5, by = 0.5),   
      expand = c(0, 0)
    ) +
    labs(
      title = title,
      subtitle = subtitle,
      y = "Responsibility",
      x = NULL,
      fill = "Follower Proportion:        "
    ) +
    theme_minimal(base_size = 20) +
    theme(
      legend.position = if (show_legend) "bottom" else "none",
      legend.spacing.x = unit(0.8, "cm"),
      legend.text = element_text(
        size = 30,
        margin = margin(l = 0.4, r = 2, unit = "cm")
      ),
      legend.title = element_text(size = 28),

      axis.text.x  = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank(),

      axis.line = element_line(color = "black", linewidth = 0.2),
      axis.ticks = element_line(color = "black"),
      axis.ticks.length = unit(0.2, "cm"),

      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),

      panel.grid.major.y = element_line(color = "grey80", linewidth = 0.5),
      panel.grid.minor.y = element_blank(),

      axis.text.y  = element_text(size = 25),
      axis.title.y = element_text(size = 25),

      plot.title = element_text(size = 35, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(
        size = 28,
        color = "grey35",
        hjust = subtitle_hjust
      )
    )

  p
}

p1 <- make_prop_barplot_raw(
  df_e1,
  "A. Experiment 1",
  subtitle = "Visible voting",
  subtitle_hjust = 0.49
)

p2 <- make_prop_barplot_raw(
  df_e2,
  "B. Experiment 2",
  subtitle = "Mechanical system",
  subtitle_hjust = 0.62,
  show_legend = TRUE
) +
  theme(
    legend.position = "bottom",
    legend.justification = "center"
  )

p3 <- make_prop_barplot_raw(
  df_e3,
  "C. Experiment 3",
  subtitle = "Advised voting",
  subtitle_hjust = 0.52
)


patched_follow_prop <- (p1 | p2 | p3) +
  plot_layout(widths = c(3, 3, 3))   

patched_follow_prop
```

```{r}
ggsave(
  filename = file.path(out_dir, "follower_proportion.pdf"),
  plot = patched_follow_prop,
  width = 25,
  height = 8,
  dpi = 300
)

```

# Proportion
```{r}
boot_mean_ci <- function(x, R = 2000, conf = 0.95) {
  x <- x[is.finite(x)]
  if (length(x) == 0) return(tibble(mean = NA_real_, ci_low = NA_real_, ci_high = NA_real_, n = 0))
  if (length(x) == 1) return(tibble(mean = mean(x), ci_low = mean(x), ci_high = mean(x), n = 1))

  boot_means <- replicate(R, mean(sample(x, replace = TRUE)))
  alpha <- (1 - conf) / 2

  tibble(
    mean = mean(x),
    ci_low = unname(quantile(boot_means, probs = alpha)),
    ci_high = unname(quantile(boot_means, probs = 1 - alpha)),
    n = length(x)
  )
}

summarise_prop_stats <- function(df, exp_label) {
  df %>%
    mutate(prop_lbl = factor(prop_lbl, levels = PROP_LEVELS)) %>%
    group_by(prop_lbl) %>%
    summarise(tmp = list(boot_mean_ci(response)), .groups = "drop") %>%
    tidyr::unnest(tmp) %>%
    mutate(exp = exp_label) %>%
    select(exp, prop_lbl, n, mean, ci_low, ci_high)
}

prop_stats_e1 <- summarise_prop_stats(df_e1, "Experiment 1")
prop_stats_e2 <- summarise_prop_stats(df_e2, "Experiment 2")
prop_stats_e3 <- summarise_prop_stats(df_e3, "Experiment 3")

prop_stats_all <- bind_rows(prop_stats_e1, prop_stats_e2, prop_stats_e3)

prop_stats_all

prop_stats_all %>%
  mutate(
    report = sprintf("M = %.2f, 95%% CI [%.2f, %.2f] (n=%d)", mean, ci_low, ci_high, n)
  ) %>%
  select(exp, prop_lbl, report)
```

```{r}
fit_proportion_exp1 = brm(
  formula = response ~ 1 + follow_proportion + (1|id),
  data = response_data_exp1 %>% filter(role=="root"),
  iter = 9000,
  seed = 1,
  file = "cache/fit_proportion_exp1",
  chains = 8
)

summary(fit_proportion_exp1)
```
```{r}
fit_proportion_exp2 = brm(
  formula = response ~ 1 + follow_proportion + (1|id),
  data = response_data_exp2 %>% filter(role=="root"),
  iter = 9000,
  seed = 1,
  file = "cache/fit_proportion_exp2",
  chains = 8
)

summary(fit_proportion_exp2)
```

```{r}
fit_proportion_exp3 = brm(
  formula = response ~ 1 + follow_proportion + (1|id),
  data = response_data_exp3 %>%
    filter(role == "mentor") %>%
    mutate(follow_proportion = as.numeric(as.character(follow_proportion)) / 100),
  iter = 9000,
  seed = 1,
  file = "cache/fit_proportion_exp3",
  chains = 8
)

summary(fit_proportion_exp3)
```


# Means and bootstrap 95% CIs by role and experiment
```{r}
ROLE_BREAKS <- c(
  "root (with knowledge)",
  "root (without knowledge)",
  "dependent",
  "independent"
)


boot_mean_ci <- function(x, conf = 0.95) {
  x <- x[!is.na(x)]
  if (length(x) == 0) return(c(lower = NA_real_, upper = NA_real_))
  ci <- Hmisc::smean.cl.boot(x, conf.int = conf)
  c(lower = unname(ci["Lower"]), upper = unname(ci["Upper"]))
}

df_exp1_min <- response_data_exp1 %>%
  filter(!is.na(role), !is.na(response)) %>%
  mutate(
    experiment = "Experiment 1",
    role_lbl = case_when(
      role == "root" & knows == TRUE  ~ "root (with knowledge)",
      role == "root" & knows == FALSE ~ "root (without knowledge)",
      role == "dependent"             ~ "dependent",
      role == "independent"           ~ "independent",
      TRUE ~ NA_character_
    ),
    role_lbl = factor(role_lbl, levels = ROLE_BREAKS)
  ) %>%
  filter(!is.na(role_lbl)) %>%
  select(experiment, role_lbl, response)

df_exp2_min <- response_data_exp2 %>%
  filter(!is.na(role), !is.na(response)) %>%
  mutate(
    experiment = "Experiment 2",
    role_lbl = case_when(
      role == "root"        ~ "root (without knowledge)",
      role == "dependent"   ~ "dependent",
      role == "independent" ~ "independent",
      TRUE ~ NA_character_
    ),
    role_lbl = factor(role_lbl, levels = ROLE_BREAKS)
  ) %>%
  filter(!is.na(role_lbl)) %>%
  select(experiment, role_lbl, response)

df_exp3_min <- response_data_exp3 %>%
  filter(!is.na(role), !is.na(response)) %>%
  mutate(
    experiment = "Experiment 3",
    role_lbl = case_when(
      role == "mentor"      ~ "root (with knowledge)",
      role == "mentee"      ~ "dependent",
      role == "independent" ~ "independent",
      TRUE ~ NA_character_
    ),
    role_lbl = factor(role_lbl, levels = ROLE_BREAKS)
  ) %>%
  filter(!is.na(role_lbl)) %>%
  select(experiment, role_lbl, response)

df_all <- bind_rows(df_exp1_min, df_exp2_min, df_exp3_min)

means_ci_tbl <- df_all %>%
  group_by(experiment, role_lbl) %>%
  summarise(
    n = sum(!is.na(response)),
    mean = mean(response, na.rm = TRUE),
    ci = list(boot_mean_ci(response)),
    .groups = "drop"
  ) %>%
  unnest_wider(ci) %>%
  rename(ci_lower = lower, ci_upper = upper) %>%
  mutate(
    mean = round(mean, 2),
    ci_lower = round(ci_lower, 2),
    ci_upper = round(ci_upper, 2)
  ) %>%
  arrange(experiment, role_lbl)

knitr::kable(
  means_ci_tbl %>% select(experiment, role_lbl, mean, ci_lower, ci_upper),
  caption = "Means and bootstrap 95% CIs by role and experiment"
)

```

# Mixed effect model for exp 1
```{r}
fit_model_one_exp1 = brm(
  formula = response ~ 1 + detailed_role + number_yes_votes + (1|id),
  
  data = response_data_exp1,
  iter = 9000,
  seed = 1,
  file = "cache/fit_model_exp1",
  chain =8
)

summary(fit_model_one_exp1)
```
```{r}
emm_role <- emmeans(
  fit_model_one_exp1,
  ~ detailed_role,
)

pairs_role<-pairs(emm_role)
pairs_role

```
# Mixed effect model for exp 2
```{r}

fit_model_one_exp2 <- brm(
  formula = response ~ 1 + role + number_yes + (1 | id),
  data = response_data_exp2,
  iter = 9000,
  seed = 1,
  file = "cache/fit_model_exp2",
  chain = 8
)

summary(fit_model_one_exp2)

```
```{r}
emm_role <- emmeans(
  fit_model_one_exp2,
  ~ role,
)

pairs_role<-pairs(emm_role)
pairs_role
```
# Mixed effect model for exp 3
```{r}
response_data_exp3 <- response_data_exp3 %>%
  mutate(
    role = factor(role),
    role = relevel(role, ref = "mentor")   
  )

fit_model_exp3 = brm(
  formula = response ~ 1 + role + outcome + n_votes + (1|id),
  data = response_data_exp3,
  iter = 9000,
  seed = 1,
  file = "cache/fit_model_exp3",
  chain =8
)

summary(fit_model_exp3)

```
```{r}
emm_role <- emmeans(
  fit_model_exp3,
  ~ role,
)

pairs_role<-pairs(emm_role)
pairs_role

```

# Exploratory Analyses
## Experiment 1
### Mixed effect model (random slopes) for exp 1
```{r}
fit_model_exp1_exploratory = brm(
  formula = response ~ 1 + detailed_role + number_yes_votes +  (1 + detailed_role + number_yes_votes| id),
  data = response_data_exp1,
  iter = 2000,
  seed = 1,
  file = "cache/fit_model_exp1_exploratory",
  chain = 6
)

summary(fit_model_exp1_exploratory)
```
```{r}
emm_role <- emmeans(
  fit_model_exp1_exploratory,
  ~ detailed_role,
)

pairs_role<-pairs(emm_role)
pairs_role

```
### follow_proportion
```{r}
fit_proportion_exp1_exploratory = brm(
  formula = response ~ 1 + follow_proportion + (1 + follow_proportion|id),
  data = response_data_exp1 %>% filter(role=="root"),
  iter = 9000,
  seed = 1,
  file = "cache/fit_proportion_exp1_exploratory",
  chains = 8
)
summary(fit_proportion_exp1_exploratory)
```
## Experiment 2
### Mixed effect model (random slopes) for exp 2
```{r}
fit_model_exp2_exploratory = brm(
  formula = response ~ 1 + role + number_yes +  (1 + role + number_yes| id),
  data = response_data_exp2,
  iter = 2000,
  seed = 1,
  file = "cache/fit_model_exp2_exploratory",
  chain = 6
)

summary(fit_model_exp2_exploratory)
```
```{r}
emm_role <- emmeans(
  fit_model_exp2_exploratory,
  ~ role,
)

pairs_role<-pairs(emm_role)
pairs_role
```
###follow_proportion
```{r}
fit_proportion_exp2_exploratory = brm(
  formula = response ~ 1 + follow_proportion + (1 + follow_proportion|id),
  data = response_data_exp2 %>% filter(role=="root"),
  iter = 9000,
  seed = 1,
  file = "cache/fit_proportion_exp2_exploratory",
  chains = 8
)
summary(fit_proportion_exp2_exploratory)
```

## Experiment 3
### Mixed effect model (random slopes) for exp 3
```{r}
fit_model_exp3_exploratory = brm(
  formula = response ~ 1 + role + outcome + n_votes + (1 + role + outcome + n_votes|id),
  data = response_data_exp3,
  iter = 2000,
  seed = 1,
  file = "cache/fit_model_exp3_exploratory",
  chain = 6
)

summary(fit_model_exp3_exploratory)
```
```{r}
emm_role <- emmeans(
  fit_model_exp3_exploratory,
  ~ role,
)

pairs_role<-pairs(emm_role)
pairs_role
```

###follow_proportion
```{r}
fit_proportion_exp3_exploratory = brm(
  formula = response ~ 1 + follow_proportion + (1 + follow_proportion|id),
  data = response_data_exp3 %>%
    filter(role == "mentor") %>%
    mutate(follow_proportion = as.numeric(as.character(follow_proportion)) / 100),
  iter = 9000,
  seed = 1,
  file = "cache/fit_proportion_exp3_exploratory",
  chains = 8
)

summary(fit_proportion_exp3_exploratory)
```
